---
title: "p8105_hw3_tem2171"
author: "Teresa Moore"
date: "2023-10-13"
output: html_document
---
## Problem 2
This problem uses the BRFSS data. DO NOT include this dataset in your local data directory; instead, load the data from the p8105.datasets package using

```{r}
library(p8105.datasets)
data("brfss_smart2010")

library(tidyverse)
library(dplyr)
library(ggplot2)
```

## Problem 2a
First, do some data cleaning:

format the data to use appropriate variable names;
focus on the “Overall Health” topic

include only responses from “Excellent” to “Poor”
organize responses as a factor taking levels ordered from “Poor” to “Excellent”

```{r Problem 2a, cleaning}
response_order <- c("Poor", "Fair", "Good", "Very good", "Excellent")

overall_health_df <- brfss_smart2010 |>
  janitor::clean_names() |>
  filter(topic == "Overall Health") |>
  transform(response = factor(response, levels = response_order, ordered = TRUE))

```


## Problem 2b
Using this dataset, do or answer the following (commenting on the results of each):

In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
#2002
states_2002 <- overall_health_df |>
  filter(year==2002) |>
  group_by(locationabbr) |>
  summarise(unique_locationdesc = n_distinct(locationdesc)) |>
  filter(unique_locationdesc >= 7)

#2010
states_2010 <- overall_health_df |>
  filter(year==2010) |>
  group_by(locationabbr) |>
  summarise(unique_locationdesc = n_distinct(locationdesc)) |>
  filter(unique_locationdesc >= 7)

```

There were `r nrow(states_2002)` states that were observed at 7 or more locations in 2002.
There were `r nrow(states_2010)` states that were observed at 7 or more locations in 2010.

## Problem 2c
Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state.

```{r}
excellent_overall_health_df <- overall_health_df |>
  filter(response == "Excellent") |>
  group_by(year, locationabbr) |>
  mutate(average_data_value = mean(data_value, na.rm = TRUE)) |>
  select(year, locationabbr, average_data_value)
```


## Problem 2d
Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
ggplot(excellent_overall_health_df, aes(x = year, y = average_data_value, group = locationabbr, color = locationabbr)) +
  geom_line() +
  labs(title = "Average Value Over Time Within Each State",
       x = "Year",
       y = "Average Data Value")
```


## Problem 2e
Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
ny_data <- overall_health_df |>
  filter(locationabbr == "NY" & year %in% c(2006, 2010)) |>
  separate(locationdesc, into = c("state", "county")) |>
  select(-state) |>
  group_by(county)

ggplot(ny_data, aes(x = data_value, y = response)) +
  geom_point(aes(color = county)) +
  facet_grid(. ~ year) +
  labs(
    title = "Distribution of Data Value for Responses in New York State",
    x = "Data Value",
    y = "Response"
  )
```

## Problem 3

Accelerometers have become an appealing alternative to self-report techniques for studying physical activity in observational studies and clinical trials, largely because of their relative objectivity. During observation periods, the devices can measure MIMS in a short period; one-minute intervals are common. Because accelerometers can be worn comfortably and unobtrusively, they produce around-the-clock observations.

This problem uses accelerometer data collected on 250 participants in the NHANES study. The participants’ demographic data can be downloaded here, and their accelerometer data can be downloaded here. Variables *MIMS are the MIMS values for each minute of a 24-hour day starting at midnight.

```{r}
demo_data = read_csv(file = "nhanes_covar.csv", skip=4) |>
  janitor::clean_names()

accel_data = read_csv(file = "nhanes_accel.csv") |>
  janitor::clean_names()
```

## Problem 3a
Load, tidy, merge, and otherwise organize the data sets. Your final dataset should include all originally observed variables; 
exclude participants less than 21 years of age
exclude those with missing demographic data
encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind).

```{r}

```


## Problem 3 b
Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. Comment on these items.

## Problem 3 c
Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, 
aggregate across minutes to create a total activity variable for each participant

Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level.
Include a trend line or a smooth to illustrate differences. 
Comment on your plot.

## Problem 3d
Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. 

Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.